---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

# Load the package from source (not the installed version) so the datasets
# table below always reflects the current working tree - this also means
# README.Rmd renders standalone (e.g. via the IDE's Preview/Render button),
# not just through `just render-readme`.
devtools::load_all(quiet = TRUE)
```

# opencodecounts

<!-- badges: start -->

[![R-CMD-check](https://github.com/bennettoxford/opencodecounts/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bennettoxford/opencodecounts/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `opencodecounts` R package provides easy access to yearly summaries of clinical code usage in England. The package makes NHS England's published datasets available in R and an interactive online tool, covering SNOMED CT codes in primary care and ICD-10/OPCS-4 codes in secondary care.

You can launch the interactive Shiny app by clicking on [*Launch Shiny app*](https://bennettoxford.github.io/opencodecounts/articles/app.html). For work in R, all available functions (including datasets) are documented in the [*R Reference*](https://bennettoxford.github.io/opencodecounts/reference/index.html) section.

## Installation

You can install the `opencodecounts` package in R with:  

```r
# install.packages("pak")
pak::pak("bennettoxford/opencodecounts")
```

## How-to guides

- [How to use the Shiny app](https://bennettoxford.github.io/opencodecounts/articles/how-to-use-shiny-app.html)
- [How to use the R package](https://bennettoxford.github.io/opencodecounts/articles/how-to-use-R-pkg.html)
- [How to use ICD-10 and OPCS-4 breakdowns](https://bennettoxford.github.io/opencodecounts/articles/how-to-use-breakdowns.html)
- [How to extract semantic tags from SNOMED CT descriptions](https://bennettoxford.github.io/opencodecounts/articles/extract-snomedct-sem-tag.html)
- [Learn about the available datasets](https://bennettoxford.github.io/opencodecounts/articles/available-datasets.html)

## Available datasets

```{r datasets-table, echo = FALSE, message = FALSE}
snomed_source <- "[Code Usage in Primary Care](https://digital.nhs.uk/data-and-information/publications/statistical/mi-snomed-code-usage-in-primary-care)"
hes_source <- "[Hospital Admitted Patient Care Activity](https://digital.nhs.uk/data-and-information/publications/statistical/hospital-admitted-patient-care-activity)"

dataset_info <- list(
  list(id = "snomed_usage", fn = "get_snomed_usage()", source = snomed_source, data = get_snomed_usage()),
  list(id = "icd10_usage", fn = "get_icd10_usage()", source = hes_source, data = get_icd10_usage()),
  list(id = "icd10_usage_breakdowns", fn = "get_icd10_usage_breakdowns()", source = hes_source, data = get_icd10_usage_breakdowns()),
  list(id = "opcs4_usage", fn = "get_opcs4_usage()", source = hes_source, data = get_opcs4_usage()),
  list(id = "opcs4_usage_breakdowns", fn = "get_opcs4_usage_breakdowns()", source = hes_source, data = get_opcs4_usage_breakdowns())
)

tidy_sources <- load_tidy_sources_config()

purrr::map_dfr(dataset_info, function(d) {
  tibble::tibble(
    Function = paste0("`", d$fn, "`"),
    Source = d$source,
    `First period` = format(min(d$data$start_date), "%Y-%m-%d"),
    `Last period` = format(max(d$data$end_date), "%Y-%m-%d"),
    `Tidy data` = paste0("[Download](", tidy_sources[[d$id]]$url, ")")
  )
}) |>
  knitr::kable(align = c("l", "l", "r", "r", "l"))
```

## For developers

See [DEVELOPERS.md](DEVELOPERS.md).
