---
title: "How to use ICD-10 and OPCS-4 breakdowns"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{how-to-use-breakdowns}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knit-options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE, message = FALSE}
library(opencodecounts)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(readr)
library(scales)
```

# Overview

Breakdowns allow you to go beyond total usage counts and explore how ICD-10 and OPCS-4 codes are used across different patient groups and clinical contexts.
For more information on all datasets available in this package see the [available datasets](available-datasets.html) vignette.

1. **Position**: Describe the count of main diagnosis (vs all dignoses) for inpatient treatment and the main procedures (vs all procedures) recorded for each available code;
2. **Sex**: female, male, and gender unknown;
3. **Age**: mostly 5-year age groups from 0 to 90 plus.

See the full list of ICD-10 breakdowns in the results below.
The same breakdowns are available for OPCS-4, but the position breakdowns are called `all_procedures` and `main_procedure` instead.

```{r list-available-breakdowns}
# Available breakdowns for ICD-10
unique(icd10_usage_breakdowns$breakdown)
```

# Worked examples

In these examples we're exploring the code usage for the ICD-10 code [F90.0 (Disturbance of activity and attention)](https://icd.who.int/browse10/2010/en#/F90.0), which is used to record Attention Deficit Hyperactivity Disorder (ADHD).
The underlying data structure looks like this:

```{r filter-adhd-code}
icd10_adhd_breakdowns <- icd10_usage_breakdowns |>
  filter(icd10_code == "F900")

icd10_adhd_breakdowns
```

## Calculating the percentage of main diagnosis coding

Calculating what proportion of usage is recorded as a main diagnosis helps distinguish between codes used to describe the primary reason for admission and those capturing co-existing conditions.

```{r calc-main-pct}
adhd_diag_positions <- icd10_adhd_breakdowns |>
  filter(breakdown %in% c("all_diagnoses", "main_diagnosis")) |>
  group_by(icd10_code, description, breakdown) |>
  summarise(n = sum(usage), .groups = "drop") |>
  pivot_wider(names_from = breakdown, values_from = n) |>
  mutate(pct_main_diagnosis = round(main_diagnosis / all_diagnoses * 100, 2))

adhd_diag_positions
```

Only `r filter(adhd_diag_positions, icd10_code == "F900")$pct_main_diagnosis`% of F90.0 coding is recorded as the main diagnosis, indicating that ADHD is typically recorded as a co-existing condition and not the main condition being treated.

## Visualising trends in sex breakdowns

```{r fig-sex-breakdowns, fig.height = 4, fig.width = 8, fig.cap = "Trends in ICD-10 code F90.0 (Disturbance of activity and attention) usage by sex in England."}
# Define all x axis breaks
x_breaks <- unique(icd10_adhd_breakdowns$end_date)

# Create figure for sex breakdowns
fig_adhd_sex_breakdowns <- icd10_adhd_breakdowns |>
  filter(breakdown %in% c("male", "female", "gender_unknown")) |> 
  ggplot(aes(x = end_date, y = usage, colour = breakdown, shape = breakdown)) +
  geom_line(alpha = .3) +  
  geom_point(size = 2) +
  scale_y_continuous(labels = label_number(accuracy = 1)) +
  scale_x_date(breaks = x_breaks, labels = label_date("%b\n%Y")) +
  scale_colour_viridis_d(end = .7, option = "C") +
  labs(x = NULL, y = "Usage count", colour = NULL, shape = NULL)

fig_adhd_sex_breakdowns
```

## Visualising trends in age breakdowns

```{r fig-age-breakdowns, fig.height = 10, fig.width = 8, fig.cap = "Trends in ICD-10 code F90.0 (Disturbance of activity and attention) usage by age group in England."}
# Define short x axis breaks (only every 4th reporting period)
x_breaks_short <- x_breaks[seq(1, length(x_breaks), by = 4)]

# Create figure for age breakdowns
fig_adhd_age_breakdowns <- icd10_adhd_breakdowns |>
  filter(str_starts(breakdown, "age")) |> 
  mutate(breakdown = reorder(breakdown, parse_number(breakdown))) |>
  ggplot(aes( x = end_date, y = usage, colour = breakdown)) +
  geom_line(alpha = .3) +  
  geom_point() +
  scale_y_continuous(labels = label_number(accuracy = 1)) +
  scale_x_date(breaks = x_breaks_short, labels = label_date("%b\n%Y")) +
  scale_colour_viridis_d(end = .75) +
  labs(x = NULL, y = "Usage count",colour = NULL, shape = NULL) +
  theme(legend.position = "none", panel.spacing = unit(.4, "cm")) +
  facet_wrap(~breakdown, ncol = 4)

fig_adhd_age_breakdowns
```
