---
title: "How to use ICD-10 and OPCS-4 breakdowns"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{how-to-use-breakdowns}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

The following code usage breakdowns are available for ICD-10 and OPCS-4:

1. Position - `main_diagnosis` and `main_procedure` describe the main reason for inpatient treatment, while co-existing conditions and other procedures are recorded under `all_diagnoses` and `all_procedures`; 

2. Sex;

3. Age - roughly 5-year age groups, with ages 0, 15, 16, 17, 18, and 19 recorded separately, and all ages over 90 aggregated.

```{r setup, message = FALSE}
library(opencodecounts)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(forcats)
library(readr)
```

```{r show-available-breakdowns}
# Available breakdowns for ICD-10
unique(icd10_usage_breakdowns$breakdown)
```

The same breakdowns are available for OPCS-4, but the position breakdowns are called `all_procedures` and `main_procedure` instead.

# Worked examples

In these examples we're looking the ICD-10 code F90.0, which is used to record ADHD (Attention Deficit Hyperactivity Disorder).

```{r filter-adhd-code}
icd10_adhd_breakdowns <- icd10_usage_breakdowns |>
  filter(icd10_code == "F900")
```

## Calculating the percentage of main diagnosis coding

```{r calc-main-pct}
adhd_diag_positions <- icd10_adhd_breakdowns |> 
  filter(breakdown %in% c("all_diagnoses", "main_diagnosis")) |> 
  group_by(icd10_code, description, breakdown) |>
  summarise(n = sum(usage), .groups = "drop") |>
  # pivot_wider(names_from = breakdown, values_from = n)
  mutate(pct = round(n / n[breakdown == "all_diagnoses"] * 100, 2))

adhd_diag_positions
```

Only 0.67% of F90.0 coding is recorded as the main diagnosis, indicating that ADHD is typically recorded as a co-existing condition and not the main condition being treated.

## Visualising trends in sex breakdowns

```{r plot-sex-breakdowns}
plot_adhd_sex_breakdowns <- icd10_adhd_breakdowns |>
  filter(breakdown %in% c("male", "female", "gender_unknown")) |> 
  ggplot(aes(
    x = end_date,
    y = usage,
    colour = breakdown,
    shape = breakdown)
  ) +
  geom_line(alpha = .3) +  
  geom_point() +
  scale_y_continuous(labels = scales::label_number(accuracy = 1)) +
  scale_x_date() +
  scale_colour_viridis_d(end = .7) +
  labs(
    x = NULL, y = "Usage count",
    colour = NULL, shape = NULL
  )

plot_adhd_sex_breakdowns
```

## Visualising trends in age breakdowns

```{r plot-age-breakdowns}
plot_adhd_age_breakdowns <- icd10_adhd_breakdowns |>
  filter(str_starts(breakdown, "age")) |> 
  mutate(breakdown = fct_reorder(breakdown, parse_number(breakdown))) |> 
  ggplot(aes(
    x = end_date,
    y = usage,
    colour = breakdown)
  ) +
  geom_line(alpha = .3) +  
  geom_point() +
  scale_y_continuous(labels = scales::label_number(accuracy = 1)) +
  scale_x_date() +
  scale_colour_viridis_d(end = .75) +
  labs(
    x = NULL, y = "Usage count",
    colour = NULL, shape = NULL
  ) +
  theme(legend.position = "none") +
  facet_wrap(~breakdown, ncol = 4)

plot_adhd_age_breakdowns
```